
# What contributes to the increasing representation of women in Moroccan Politics? A time series model

The full written analysis is available on the pdf file.

```{r setup, include=FALSE}
# Formatting options and packages
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE,
                      error = FALSE, 
                      message = FALSE, 
                      fig.align='center', fig.width = 5.0, fig.height = 3) #just for the outline

library(formatR)                # allows to format R code
library(tidyverse)              # allows for Tidy Verse data-cleaning commands

# Paper specific packages
library(foreign)                # package allows R to read Stata datasets ("read_dta" command)
library(ggcorrplot)             # package for correlation matrix plot ("ggcorrplot" command)
library(psych)                  # package for data analysis ("describe" command)
library(forecast)               # package needed for time-series analysis
library(car)
library(prais)
library(orcutt)
library(ecm)
library(sandwich)

# Set-up
set.seed(20136)                 #important because allows to trace random procedures 
rm(list = ls(all = TRUE))       #remove all previous objects 
```
## Data Exploration

#### Data Upload and Cleaning

Variables List: 
    pr: percentage of parliamentary seats held by women, a continuous variables with possible values going from 0 to 100%. 
    lfpr: percentage of women in the workforce, a continuous variable with possible values from 0 to 100%. 
    rate: fertility rate, which is the number of children per woman ages 15 to 50, to measure motherhood. 
    gnipc: Gross National Income per capita of women in US dollars, as a proxy to women's wealth in the country.The value is in current dollars.
    eys: expected years of schooling for women in Morocco
    fdi: Foreign Direct Investment flows, in current US dollars, in billions. As a proxy for country’s increased integration in globalization
    growth: GDP growth rate (percentage) as a proxy to economic development 
    exp: freedom of expression sub-index by V-dem instute, as a proxy to the strength of the democratic region. Ranges from 0 to 1.

```{r}
# Uploading World Bank Data
brut <- read.csv("HDR21-22_Composite_indices_complete_time_series.csv", header = TRUE)

# Focus on Morocco
morocco <- brut[brut$country == "Morocco",]

# Keep only the variables of interest
interest <- c("le_","eys_f","gni_pc_f","co2_prod_","pr_f","lfpr_f")
options <- paste0("^(", paste(interest, collapse = "|"), ")") # to use in grepl() and check for options

names <- colnames(morocco)
index <- c() # to keep track of which columns to keep

for(i in 1:length(names)){
    if(grepl(options, names[i])){index <- c(index,i)}
}

morocco <- morocco[,index]

# Turn the variables into columns with the years in rows
names <- colnames(morocco)

  # Expected Years of schooling
index <- c()
for(i in 1:length(names)){
    if(grepl("^eys", names[i])){index <- c(index,i)}
}
    # Keeping only from 1990 to 2021
eys <- unlist(unname(as.vector(morocco[,index[1:32]])))

  # Gross national income per capita
index <- c()
for(i in 1:length(names)){
    if(grepl("^gni_pc_f", names[i])){index <- c(index,i)}
}
    # Keeping only from 1990 to 2021
gnipc <- unlist(unname(as.vector(morocco[,index[1:32]])))

  # Women in Parliament 
index <- c()
for(i in 1:length(names)){
    if(grepl("^pr_f", names[i])){index <- c(index,i)}
}
    # Keeping only from 1990 to 2021
pr <- unlist(unname(as.vector(morocco[,index[1:32]])))

  # Women in Labor Force
index <- c()
for(i in 1:length(names)){
    if(grepl("^lfpr_f", names[i])){index <- c(index,i)}
}
    # Keeping only from 1990 to 2021
lfpr <- unlist(unname(as.vector(morocco[,index[1:32]])))

# Fertility rate
fertility <- read.csv("morocco-population-2023-04-27.csv", sep = ",", header = T)
fertility$date <- 1950:2100

    # Keeping only from 1990 to 2021
min <- which(fertility$date == 1990)
max <- which(fertility$date == 2021)
rate <- fertility$Births.per.Woman[min:max]

# Foreign Direct Investment
investment <- read.csv("morocco-foreign-direct-investment.csv", sep = ",", header = T)
investment$date <- 1970:2021

    # Keeping only from 1990 to 2021
min <- which(investment$date == 1990)
fdi <- investment$Inflows[min:nrow(investment)]

# GDP Growth rate
economic <- read.csv("morocco-economic-growth-rate.csv", sep = ",", header = T)
economic$date <- 1960:2021

  # Keeping only from 1990 to 2021
min <- which(economic$date == 1990)
growth <- economic$Annual...Change[min:nrow(economic)]

# Freedom of Expression
expression <- read.csv("democracy.csv", sep = ",", header = T)
  
  # Keeping only Morocco from 1990 to 2021
expression <- expression[expression$Entity == "Morocco",]
min <- which(expression$Year == 1990)
max <- which(expression$Year == 2021)
exp <- expression$freeexpr_vdem_owid[min:max]

# Final dataset
maroc <- data.frame(pr, growth, fdi, exp, rate, lfpr, gnipc, eys)
maroc$years <- 1990:2021

# Turning all variables into time series model
  # Non time series
maroc2 <- maroc
  # Time series
for(i in 1:ncol(maroc)){
  maroc[,i] <- ts(maroc[,i], start = 1990, frequency = 1)
}

rm(list = setdiff(ls(), c("maroc","maroc2")))

```

#### Data Frequency and Description

##### Table 1 - Frequency Distributions 

```{r}
sum_pr <- unname(summary(maroc$pr))
sum_growth <- unname(summary(maroc$growth))
sum_fdi <- unname(summary(maroc$fdi))
sum_exp <- unname(summary(maroc$exp))
sum_rate <- unname(summary(maroc$rate))
sum_lfpr <- unname(summary(maroc$lfpr))
sum_gnipc <- unname(summary(maroc$gnipc))
sum_eys <- unname(summary(maroc$eys))

summary <- data.frame(n = rep(nrow(maroc),8),
                      min = rep(NA,8),
                      Q1 = rep(NA,8),
                      median = rep(NA,8),
                      mean = rep(NA,8),
                      Q3 = rep(NA,8),
                      max = rep(NA,8))
rownames(summary) <- c("pr", "growth", "fdi", "exp", "rate", "lfpr", "gnipc", "eys")

values <- rbind(sum_pr,sum_growth,sum_fdi,sum_exp,sum_rate,sum_lfpr,sum_gnipc,sum_eys)
summary[,2:7] <- values
summary
```

##### Table 2 - Plotting variables over time

```{r}
plotdfts <- maroc %>%
  dplyr::select(pr, growth, fdi, exp, rate, lfpr, gnipc, eys, years) %>%
  gather(key = "variable", value = "value", -years)

# Table 2.A - Plot all variables over time
ggplot(plotdfts, aes(x = years, y = value)) + 
  geom_line(aes(color = variable), size = 1) +
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#10bb00", "#bb1c00", "#bb6d00", "#bb9f00", "#0022bb", "#b200bb")) +
  theme_minimal() + labs(title = "Independent Variable and Dependent Variable over time")

# Because of the large value of the gnipc data, its scale does not allow to see changes in the other variables. Instead, I will plot all the variables over time but not gnipc.
plotdfts <- maroc %>%
  dplyr::select(pr, growth, fdi, exp, rate, lfpr, eys, years) %>%
  gather(key = "variable", value = "value", -years)

# Table 2.B - Plot all variables over time excpet for gnipc
ggplot(plotdfts, aes(x = years, y = value)) + 
  geom_line(aes(color = variable), size = 1) +
  scale_color_manual(values = c("#00AFBB", "#E7B800", "#10bb00", "#bb6d00", "#bb9f00", "#0022bb", "#b200bb")) +
  theme_minimal()+ labs(title = "Independent Variable and Dependent Variable over time without gnipc")
```

#### Table 3 - Correlation Matrix

```{r}
dfts2 <- maroc[1:(ncol(maroc)-1)]

corr <- round(cor(dfts2, use = "complete.obs"), 2)

ggcorrplot(corr, type = "full", lab = TRUE,
           outline.col = "white",
           ggtheme = ggplot2::theme_gray,
           colors = c("#E46726", "white", "#6D9EC1"), 
           lab_col = "black", lab_size = 2, 
           tl.cex = 8, tl.col = "black")

rm(list = setdiff(ls(), c("maroc","maroc2")))
```

## Data Analysis

### Regression Analysis and Serial Correlation

#### Table 4 - OLS Regression Analysis & Serial Correlation

##### Table 4.A - OLS Regression Model

```{r}
mod1 <- lm(pr ~ growth + fdi + exp + rate + lfpr + gnipc + eys, data = maroc)
summary(mod1)
```

It is quite common for time series data to not exhibit statistical significance under the OLS regression model. This is because time series data often contains autocorrelation (also known as serial correlation), which violates the assumption of independent and identically distributed errors in the OLS regression model.

The coefficient estimates for lfpr and eys are statistically significant.

##### Table 4.B - Derbin-Watson test

```{r}
dwtest(mod1)
```

The Durbin-Watson test statistic is of 1.3981. This value is under the Lowerbound for 32 observations and 7 explanatory variables. Therefore, we have enough evidence to reject the null hypothesis. There is evidence of serial correlation.

#### Table 5 - Plotting explanotary variables against response variable to check for serial correlation

```{r}
# Scatterplots of the dependent variables against the response variables
plot(maroc2$growth, maroc2$pr)
plot(maroc2$fdi, maroc2$pr)
plot(maroc2$exp, maroc2$pr)
plot(maroc2$rate, maroc2$pr)
plot(maroc2$lfpr, maroc2$pr)
plot(maroc2$gnipc, maroc2$pr)
plot(maroc2$eys, maroc2$pr)
```

The scatterplots show that the following variables appear not to show a somewhat linear form: fdi, growth, and lfpr.

### Testing for Non-Stationarity of the variables

#### Table 7 - Plotting variables against time to check for non-stationarity

```{r}
# Table 7.A - Change in Percentage of Women Holding Parliamentary Seats in Morocco over Time
ggplot(maroc, aes(x = years, y = pr)) + 
  geom_line() + labs(title = "Change in Percentage of Women Holding Parliamentary Seats in Morocco over Time")

# Table 7.B - Change in GDP growth rate in Morocco over Time
ggplot(maroc, aes(x = years, y = growth)) + 
  geom_line() + labs(title = "Change in GDP growth rate in Morocco over Time")

# Table 7.C - Change in FDI inflow in Morocco over Time
ggplot(maroc, aes(x = years, y = fdi)) + 
  geom_line() + labs(title = "Change in FDI inflow in Morocco over Time")

# Table 7.D - Change in the Freedom of Expression score in Morocco over Time
ggplot(maroc, aes(x = years, y = exp)) + 
  geom_line() + labs(title = "Change in the Freedom of Expression score in Morocco over Time")

# Table 7.E - Change in fertility rate in Morocco over Time
ggplot(maroc, aes(x = years, y = rate)) + 
  geom_line() + labs(title = "Change in fertility rate in Morocco over Time")

# Table 7.F - Change in Percentage of Women in the workforce in Morocco over Time
ggplot(maroc, aes(x = years, y = lfpr)) + 
  geom_line() + labs(title = "Change in Percentage of Women in the workforce in Morocco over Time")

# Table 7.G - Change in the Gross National Income per capita of women in Morocco over Time 
ggplot(maroc, aes(x = years, y = gnipc)) + 
  geom_line() + labs(title = "Change in the Gross National Income per capita of women in Morocco over Time")
 
# Table 7.H - Change in expected years of schooling for women in Morocco over Time
ggplot(maroc, aes(x = years, y = eys)) + 
  geom_line() + labs(title = "Change in expected years of schooling for women in Morocco over Time")

```

There is strong evidence of non-stationarity for most of the variables.

#### Table 8 - Checking for stationarity using Autocorrelation Function

```{r}
Acf(maroc$pr)
Acf(maroc$growth)
Acf(maroc$fdi)
Acf(maroc$exp)
Acf(maroc$rate)
Acf(maroc$lfpr)
Acf(maroc$gnipc)
Acf(maroc$eys)
```

#### Table 9 - Checking for stationarity using the Augmmented Dicky-Fueller test

```{r}
library(tseries)
dfts <- maroc
# Creating lagged variables
dfts$growthlag  <- sapply(1:nrow(dfts), function(x) as.numeric(dfts$growth[x-1]))
dfts$fdilag  <- sapply(1:nrow(dfts), function(x) dfts$fdi[x-1])
dfts$explag  <- sapply(1:nrow(dfts), function(x) dfts$exp[x-1])
dfts$ratelag  <- sapply(1:nrow(dfts), function(x) dfts$rate[x-1])
dfts$lfprlag  <- sapply(1:nrow(dfts), function(x) dfts$lfpr[x-1])
dfts$gnipclag  <- sapply(1:nrow(dfts), function(x) dfts$gnipc[x-1])
dfts$eyslag  <- sapply(1:nrow(dfts), function(x) dfts$eys[x-1])

# turn the value for the first year into NA, as there is no lag it
dfts[1,10:16]<- NA
for (i in 1:ncol(dfts)) {
  dfts[, i] <- as.numeric(dfts[, i])
}

# Create first differences
dfts$diffgrowth <- dfts$growth-dfts$growthlag
dfts$difffdi <- dfts$fdi-dfts$fdilag
dfts$diffexp <- dfts$exp-dfts$explag
dfts$diffrate <- dfts$rate-dfts$ratelag
dfts$difflfpr <- dfts$lfpr-dfts$lfprlag
dfts$diffgnipc <- dfts$gnipc-dfts$gnipclag
dfts$diffeys <- dfts$eys-dfts$eyslag
```

Table 9.A - growth
```{r}
  # (a) Regress first difference on the lagged variable
  mod1 <- lm(diffgrowth ~ growthlag, data = dfts)
  summary(mod1)

    # Durbin-Watson statistic
    durbinWatsonTest(mod1)

  # (b) Regress first difference on the lagged variable and the time variable
  mod2 <- lm(diffgrowth ~ growthlag + years, data = dfts)
  summary(mod2)
     
    # Durbin-Watson statistic
    durbinWatsonTest(mod2)
  
  # (c) Formal test: Augment Dicky-Fueller
    adf.test(maroc$growth)
```

Table 9.B - fdi
```{r}
  # (a) Regress first difference on the lagged variable
  mod1 <- lm(difffdi ~ fdilag, data = dfts)
  summary(mod1)

    # Durbin-Watson statistic
    durbinWatsonTest(mod1)

  # (b) Regress first difference on the lagged variable and the time variable
  mod2 <- lm(difffdi ~ fdilag + years, data = dfts)
  summary(mod2)
     
    # Durbin-Watson statistic
    durbinWatsonTest(mod2)
  
  # (c) Formal test: Augment Dicky-Fueller
    adf.test(maroc$fdi)
```

Table 9.C - exp
```{r}
  # (a) Regress first difference on the lagged variable
  mod1 <- lm(diffexp ~ explag, data = dfts)
  summary(mod1)

    # Durbin-Watson statistic
    durbinWatsonTest(mod1)

  # (b) Regress first difference on the lagged variable and the time variable
  mod2 <- lm(diffexp ~ explag + years, data = dfts)
  summary(mod2)
     
    # Durbin-Watson statistic
    durbinWatsonTest(mod2)
  
  # (c) Formal test: Augment Dicky-Fueller
    adf.test(maroc$exp)
```

Table 9.D - rate
```{r}
  # (a) Regress first difference on the lagged variable
  mod1 <- lm(diffrate ~ ratelag, data = dfts)
  summary(mod1)

    # Durbin-Watson statistic
    durbinWatsonTest(mod1)

  # (b) Regress first difference on the lagged variable and the time variable
  mod2 <- lm(diffrate ~ ratelag + years, data = dfts)
  summary(mod2)
     
    # Durbin-Watson statistic
    durbinWatsonTest(mod2)
  
  # (c) Formal test: Augment Dicky-Fueller
    adf.test(maroc$rate)
```

Table 9.E - lfpr
```{r}
  # (a) Regress first difference on the lagged variable
  mod1 <- lm(difflfpr ~ lfprlag, data = dfts)
  summary(mod1)

    # Durbin-Watson statistic
    durbinWatsonTest(mod1)

  # (b) Regress first difference on the lagged variable and the time variable
  mod2 <- lm(difflfpr ~ lfprlag + years, data = dfts)
  summary(mod2)
     
    # Durbin-Watson statistic
    durbinWatsonTest(mod2)
  
  # (c) Formal test: Augment Dicky-Fueller
    adf.test(maroc$lfpr)
```

Table 9.F - gnipc
```{r}
  # (a) Regress first difference on the lagged variable
  mod1 <- lm(diffgnipc ~ gnipclag, data = dfts)
  summary(mod1)

    # Durbin-Watson statistic
    durbinWatsonTest(mod1)

  # (b) Regress first difference on the lagged variable and the time variable
  mod2 <- lm(diffgnipc ~ gnipclag + years, data = dfts)
  summary(mod2)
     
    # Durbin-Watson statistic
    durbinWatsonTest(mod2)
  
  # (c) Formal test: Augment Dicky-Fueller
    adf.test(maroc$gnipc)
```

Table 9.G - eys
```{r}
  # (a) Regress first difference on the lagged variable
  mod1 <- lm(diffeys ~ eyslag, data = dfts)
  summary(mod1)

    # Durbin-Watson statistic
    durbinWatsonTest(mod1)

  # (b) Regress first difference on the lagged variable and the time variable
  mod2 <- lm(diffeys ~ eyslag + years, data = dfts)
  summary(mod2)
     
    # Durbin-Watson statistic
    durbinWatsonTest(mod2)
  
  # (c) Formal test: Augment Dicky-Fueller
    adf.test(maroc$eys)
```

All my variables are non-stationary when tested using the Augmented Dicky-Fueller test, so I need to test for cointegration.

### Table 10 - Testing for Co-Integration between the variables

```{r}
# Residuals of original regression
mod4 <- lm(pr ~ growth + fdi + exp + rate + lfpr + gnipc + eys, data = maroc)
dfts <- maroc
dfts$res2 <- residuals(mod4)

# Create first difference of residuals
dfts$res2lag <- sapply(1:nrow(dfts), function(x) dfts$res2[x-1])
dfts$res2lag <- car::recode(as.numeric(dfts$res2lag), "numeric(0)=NA")
dfts$difres2 <- dfts$res2-dfts$res2lag

# Create lagged first difference of residuals
dfts$lagdifres2 <- sapply(1:nrow(dfts), function(x) dfts$difres2[x-1])
dfts$lagdifres2 <- car::recode(as.numeric(dfts$lagdifres2), "numeric(0)=NA")
```

##### Table 10.A - Testing for Co-Integration using the first difference of residuals

```{r}
mod5 <- lm(difres2 ~ res2lag, data = dfts)
summary(mod5)

# Dicky-Fueller Test
durbinWatsonTest(mod5)
```

##### Table 10.B - Testing for Co-Integration, adding the time variable

```{r}
mod6 <- lm(difres2 ~ res2lag + years, data = dfts)
summary(mod6)

# Dicky-Fueller Test
durbinWatsonTest(mod6)
```

##### Table 10.C - Testing for Co-Integration, adding the lagged first difference of variables

```{r}
mod7 <- lm(difres2 ~ res2lag + years + lagdifres2, data = dfts)
summary(mod7)

# Dicky-Fueller Test
durbinWatsonTest(mod7)
```

### Table 11 - Correcting for Autocorrelation in the Original OLS

```{r}
pw <- prais_winsten(pr ~ growth + fdi + exp + rate + lfpr + gnipc + eys, index = 1, data = maroc)
summary(pw)
```

### Table 12 - Difference in Differences

```{r}
# Dummy variable to indicate treatment
dfts$treated <- ifelse(dfts$years>=2004,1,0)

# DID estimator
did <- lm(pr ~ (growth + fdi + exp + rate + lfpr + gnipc + eys)*treated, data = dfts)
summary(did)
```

